---
title: "Rally 6A Study Inventory"
author: "Andrew Mertens"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=F}

#-----------------------------------
# setup
#
# input files:
#    heatmap_df.Rdata
#-----------------------------------

#-----------------------------------
# preamble
#-----------------------------------
rm(list=ls())
library('knitr')
library('tidyverse')
library('stringr')
library('scales')
library('RColorBrewer')
library('gridExtra')


# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"


#hbgdki pallet
tableau10 <- c("#1F77B4","#FF7F0E","#2CA02C","#D62728",
  "#9467BD","#8C564B","#E377C2","#7F7F7F","#BCBD22","#17BECF")

#define a color for fonts
textcol <- "grey20"

#Plot theme
theme_set(theme_bw())


#-----------------------------------
# load the monthly summary data
#-----------------------------------

load('C:/Users/andre/Documents/Rally-6/Results/heatmap_df.Rdata')

#Drop non-India cohorts from multi-site studies:
df <-df[df$COUNTRY=="INDIA",]

#drop obs. of children over 4 years old
df <- df[as.numeric(as.character(df$agemonth))<48,]

#Tablulate number of months and overall mean anthro for each study
tb<-df %>% group_by(STUDYID) %>% summarize(Nmonths=n(), aveWHZ=mean(meanWHZ, na.rm=T))

df <- left_join(df, tb, by="STUDYID")

#remove grant identifiers
df$STUDYID<- gsub("^k.*?-" , "", df$STUDYID)
df$STUDYID <- factor(df$STUDYID, levels=unique(df$STUDYID))

```


##List of studies on GHAP from INDIA

  *  Mal-ED - India Cohort
  *  COHORTS - India Cohort
  *  ki1000301 - DIVIDS
  *  ki1000108 - IRC
  *  ki1000108 - TDC
  *  ki1000108 - CMC-V-BCS-2002
  *  ki1000304 - VITAMIN-A
  *  ki1000304 - NEOVITA
  *  ki1000304 - Vitamin-B12
  *  ki1000304 - ZnMort
  *  ki1000304 - EU
  *  ki1000304 - LBW
  *  ki1000304b - SAS-FoodSuppl
  *  ki1000304b - SAS-CompFeed
  *  ki1000304b - ZincInf
  *  ki1000304c - IMNCI
  *  ki1000302 - TTBW
  *  ki1000306 - ZincSGA
  *  ki1000303 - BtS
  *  ki1000303 - NBSP
  *  ki1000305 - CordBloodCytometry
  *  GEMS-1
  *  GEMS-1A
  *  ANISA
  *  DSS
  *  MAHANTrust
  


### Study exclusion 

Drop cross-sectional studies and studies without longitudinal anthropometry data


```{r, echo=F}
c(as.character(df$STUDYID[df$Nmonths==1]), "GEMS-1","GEMS-1A","ANISA", "DSS")

df <- df[df$Nmonths>1,]
```

Drop studies without anthropometry

```{r, echo=F}
c(as.character(df$STUDYID[is.na(df$aveWHZ)]),"MAHANTrust")

df <- df[!(is.na(df$aveWHZ)),]
```

Drop studies only measuring children older than 2 years
```{r, echo=F}
"BTS"
```


### Study descriptions 


List descriptions of longitudinal studies with anthropometry data from India:

```{r, echo=F}

load('C:/Users/andre/Documents/Rally-6/Results/metadata6A.Rdata')

#Drop studies without anthro or cross-sectional studies from the metadata

meta<-meta[meta$study_id %in% df$STUDYID,c(2,7)]

knitr::kable(meta)

```


Note that the ZincSGA ("Zinc supplementation in infants born small for gestational age") and Vitamin B12 ("Vitamin B12 supplementation") seem to be missing from the get_sudylist_anthro() metadata list 







Note that the following studies enrol acutely ill children:

  *  ZincSGA -small for gestational age
  *  DIVIDS -low birth weight
  *  LBW -low birth weight
  *  ZincInf -children with diarrhea
  
  
The followinmg studies did not start measurement at birth and therefore misses a key risk period for wasting / stunting.

  *  ki1000304 -EU (6 months)
  *  ki1000304b -SAS-FoodSuppl (4 months)
  *  ki1000304 -Vitamin-B12 (6 months)
  *  ki1000304b -ZincInf (anytime between 1-5 months)
  *  ki100030 -ZnMort (anytime between 1-23 months)
  



```{r, echo=F}
#Order studies by age range of children
df <- df %>% arrange(-Nmonths)

#Drop missing study
df <- df %>% filter(!is.na(STUDYID))




#Make N categories
df$Ncat <- cut(df$N,breaks=c(0,10,50,150,250,500,1000,45574),labels=c("<10","10-50","50-150","150-250","250-500","500-1000",">1000"))
df$Ncat <- factor(df$Ncat)

#Set mean anthro to NA when less than 10 observations
df$meanWHZ[df$Ncat=="<10"] <- NA
df$meanHAZ[df$Ncat=="<10"] <- NA
df$wast[df$Ncat=="<10"] <- NA
df$stunt[df$Ncat=="<10"] <- NA
df$sevwast[df$Ncat=="<10"] <- NA
df$sevstunt[df$Ncat=="<10"] <- NA


#Make anthro categories
df$Ncat <- cut(df$N,breaks=c(0,10,50,150,250,500,1000,45574),labels=c("<10","10-50","50-150","150-250","250-500","500-1000",">1000"))
df$Ncat <- factor(df$Ncat)


df$meanWHZ<- cut(df$meanWHZ,breaks=c(-10,-3,-2.5,-2,-1.5,-1.25,-1,-0.75,-0.5,0,10),labels=c("< -3 ","[-3, -2.5)","[-2.5,-2)","[-2, -1,5)","[-1.5, -1.25)","[-1.25, -1)","[-1, -0.75)","[-0.75,-0.5)","[-0.5, 0)",">0"))
df$meanHAZ<- cut(df$meanHAZ,breaks=c(-10,-3,-2.5,-2,-1.5,-1.25,-1,-0.75,-0.5,10),labels=c("< -3 ","[-3, -2.5)","[-2.5,-2)","[-2, -1,5)","[-1.5, -1.25)","[-1.25, -1)","[-1, -0.75)","[-0.75,-0.5)",">-0.5"))
df$wast<- cut(df$wast,breaks=c(0,10,20,30,40,50,75,100),labels=c("0-10","10-20","20-30","30-40","40-50","50-75",">75"))
df$stunt<-  cut(df$stunt,breaks=c(0,10,20,30,40,50,75,100),labels=c("0-10","10-20","20-30","30-40","40-50","50-75",">75"))
df$sevwast<-  cut(df$sevwast,breaks=c(0,2.5,5,10,15,20,30,100),labels=c("0-2.5","2.5-5","5-10","10-15","15-20","20-30",">30"))
df$sevstunt<-  cut(df$sevstunt,breaks=c(0,5,10,15,20,30,50,100),labels=c("0-5","5-10","10-15","15-20","20-30","30-50", ">50"))

# heat map plot scheme
hm <- ggplot(df,aes(x=agemonth, y=STUDYID)) +
  geom_tile(colour="white",size=0.25)+
  coord_equal()+
  theme_bw() +
  theme(
    legend.title=element_text(color=textcol,size=8),
    legend.margin = margin(grid::unit(0.1,"cm")),
    legend.text=element_text(colour=textcol,size=7,face="bold"),
    legend.key.height=grid::unit(0.2,"cm"),
    legend.key.width=grid::unit(1,"cm"),
    legend.position = "bottom",
    axis.text.x=element_text(size=8,colour=textcol,angle=0,vjust=0.5),
    axis.text.y=element_text(size=8,vjust = 0.2,colour=textcol),
    axis.ticks=element_line(size=0.4),
    plot.title=element_text(colour=textcol,hjust=0,size=12,face="bold"),
    strip.text.x = element_text(size=10),
    strip.text.y = element_text(angle=270,size=10),
    plot.background=element_blank(),
    panel.border=element_blank())

```




### Heatmap of number of observations per month of child age by study

Heatmap displays number of observations within each childage in months for each study with longitudinal anthropometry data for the first 48 months (4 years) of a child's life.

```{r, echo=F}
N_hm <- hm +
  aes(fill=Ncat) +
  labs(x="Child age in months",y="",title="Number observations by study and child age") +
   scale_fill_brewer(palette = "Greens",na.value="grey90",
                    guide=guide_legend(title="N obs.",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))

w <- 10
h <- 7
ggsave("Nheatmap.pdf", N_hm, width = w, height = h, units = "in")
N_hm
```


### Heatmap of mean WHZ

```{r, echo=F}
#Remake heatmap, subset to healthy cohorts and children age 0-24 months
d <- df %>% mutate(agemonth=as.numeric(as.character(agemonth))) %>%  
  filter(agemonth < 25) %>% 
  filter(STUDYID!="DIVIDS" & STUDYID!="LBW" & 
           STUDYID!="ZincSGA" & STUDYID!="ZincInf")

hm <- ggplot(d,aes(x=agemonth, y=STUDYID)) +
  geom_tile(colour="white",size=0.25)+
  coord_equal()+
  theme_bw() +
  theme(
    legend.title=element_text(color=textcol,size=8),
    legend.margin = margin(grid::unit(0.1,"cm")),
    legend.text=element_text(colour=textcol,size=7,face="bold"),
    legend.key.height=grid::unit(0.2,"cm"),
    legend.key.width=grid::unit(1,"cm"),
    legend.position = "bottom",
    axis.text.x=element_text(size=8,colour=textcol,angle=0,vjust=0.5),
    axis.text.y=element_text(size=8,vjust = 0.2,colour=textcol),
    axis.ticks=element_line(size=0.4),
    plot.title=element_text(colour=textcol,hjust=0,size=12,face="bold"),
    strip.text.x = element_text(size=10),
    strip.text.y = element_text(angle=270,size=10),
    plot.background=element_blank(),
    panel.border=element_blank())

whzhm <- hm +
  aes(fill=meanWHZ) +
  labs(x="Child age in months",y="",title="Mean WHZ from birth to 24 months among enrolled-healthy cohorts") +
  #scale_fill_gradient(low = tableau10[2], high = tableau10[3],na.value="grey90",
   scale_fill_brewer(palette = "Oranges",na.value="grey90", direction=-1,
                    guide=guide_legend(title="WHZ",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
 
whzhm
ggsave("WHZheatmap.pdf", whzhm, width = w, height = h, units = "in")

head(d)
```

Cells are colored grey is there are less than 10 observations in a cell. Note smaller mean WHZ early in children's lives in some cohorts, and around 1-2 years in others.

### Heatmap of mean HAZ

```{r, echo=F}
hazhm <- hm +
  aes(fill=meanHAZ) +
  labs(x="Child age in months",y="",title="Mean HAZ from birth to 24 months among enrolled-healthy cohorts ") +
  #scale_fill_gradient(low = tableau10[2], high = tableau10[10],na.value="grey90",
  scale_fill_brewer(palette = "Blues",na.value="grey90", direction=-1,
                    guide=guide_legend(title="HAZ",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
 
hazhm
ggsave("HAZheatmap.pdf", hazhm, width = w, height = h, units = "in")

```

Note smaller mean HAZ as the child approaches 2 years in age.

### Heatmap of wasting prevalence

```{r, echo=F}
#CATEGORIZE
wasthm <- hm +
  aes(fill=wast) +
  labs(x="Child age in months",y="",title="Wasting Prevalence from birth to 24 months among enrolled-healthy cohorts") +
  #scale_fill_gradient(low = tableau10[3], high = tableau10[2],na.value="grey90",
     scale_fill_brewer(palette = "Oranges",na.value="grey90",
                    guide=guide_legend(title="Prevalence",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
ggsave("Wastheatmap.pdf", wasthm, width = w, height = h, units = "in")

wasthm
```

Earlier peak in wasting prevalence

### Heatmap of stunting prevalence

```{r, echo=F}
stunthm <- hm +
  aes(fill=stunt) +
  labs(x="Child age in months",y="",title="Stunting Prevalence from birth to 24 months among enrolled-healthy cohorts ") +
  #scale_fill_gradient(low = tableau10[10], high = tableau10[2],na.value="grey90",
  scale_fill_brewer(palette = "Blues",na.value="grey90",
                    guide=guide_legend(title="Prevalence",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
 ggsave("stuntheatmap.pdf", stunthm, width = w, height = h, units = "in")

stunthm
```

Increased faltering to stunting as the child ages.

### Heatmap of severe wasting prevalence

```{r, echo=F}
#CATEGORIZE
sevwasthm <- hm +
  aes(fill=sevwast) +
  labs(x="Child age in months",y="",title="Severe Wasting Prevalence from birth to 24 months among enrolled-healthy cohorts") +
     scale_fill_brewer(palette = "Oranges",na.value="grey90",
                    guide=guide_legend(title="Prevalence",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
ggsave("sevwastheatmap.pdf", sevwasthm, width = w, height = h, units = "in")

sevwasthm
```

### Heatmap of severe stunting prevalence

```{r, echo=F}
sevstunthm <- hm +
  aes(fill=sevstunt) +
  labs(x="Child age in months",y="",title="Severe Stunting Prevalence from birth to 24 months among enrolled-healthy cohorts ") +
  scale_fill_brewer(palette = "Blues",na.value="grey90",
                    guide=guide_legend(title="Prevalence",title.vjust = 1,
                                       label.position="bottom",label.hjust=0.5,nrow=1))
ggsave("sevstuntheatmap.pdf", sevstunthm, width = w, height = h, units = "in")
 
sevstunthm
```

Much more severe stunting int ZincSGA, a study that enrolled small for gestational age infants. There doesn not appear to be catch-up growth on average in these children.









